ProofOfRho

### 2020-12-30 SIR_manuscript.tex

\newgeometry{margin=1cm} % modify this if you need even more space
\begin{landscape}
\begin{table}[htp]
\centering
{\tiny %
\begin{tabular}{|c|ccc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of onward flow from the awaiting to reported or untested compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
  $\eta_w$  & Relative probability of transmission for isolated awaiting individuals & - & - \\ \hline
  $\eta_c$  & Relative probability of transmission for isolated confirmed individuals & - & -  \\ \hline
  $\Lambda$ & Force of infection & 1/time & - \\ \hline
  $p_S$ & Probability of false positive for susceptible& - & 0 \\ \hline
  $p_I$ & Probability of being infected and tested positive & - & 1 \\ \hline
  $p_R$ & Probability of being recovered and tested positive & - & 0.5 \\ \hline
  $W_S, W_I, W_R$ & Relative testing weight & - & 
  \begin{minipage}[t]{0.21\columnwidth}%
 Random testing: $W_S=W_I=W_R=1$, \\Non-random testing: $W_S=0.3, W_I=W_R=1$
\end{minipage} \\ \hline

>>>>

\newgeometry{margin=1cm} % modify this if you need even more space
\begin{landscape}
\begin{table}[htp]
\centering
{\tiny %
\begin{tabular}{|c|ccc|} \hline
  Symbol & Description & Unit & Value \\ \hline
  $N_0$     & Total population size & people & $10^6$ \\ \hline
  $\omega$  & Rate of onward flow from the awaiting to reported or untested compartments  & 1/day & - \\ \hline
  $\gamma$ & Recovery rate & 1/day & 1/3 \\ \hline 
  $\rho$   & Per capita testing intensity & 1/day & 0.01 \\ \hline 
  $\eta_w$  & Relative probability of transmission for isolated awaiting individuals & - & - \\ \hline
  $\eta_c$  & Relative probability of transmission for isolated confirmed individuals & - & -  \\ \hline
  $\Lambda$ & Force of infection & 1/day & - \\ \hline
  $p_S$ & Probability of false positive for susceptible& - & 0 \\ \hline
  $p_I$ & Probability of being infected and tested positive & - & 1 \\ \hline
  $p_R$ & Probability of being recovered and tested positive & - & 0.5 \\ \hline
  $W_S, W_I, W_R$ & Relative testing weight & - & 
  \begin{minipage}[t]{0.21\columnwidth}%
 Random testing: $W_S=W_I=W_R=1$, \\Non-random testing: $W_S=0.3, W_I=W_R=1$
\end{minipage} \\ \hline

### 2020-12-30 codes/sir_macpan.Rnw
seperating functions

<<install the package, echo=F  >>=
# install remotes package if necessary:
while (!require(remotes)) {
    install.packages("remotes")
}
## install development version of bbmle:
if (!require("bbmle") || packageVersion("bbmle") < "1.0.23.5") {
    remotes::install_github("bbolker/bbmle")
}
## install the target package and all its dependencies:
remotes::install_github("bbolker/McMasterPandemic",
                        dependencies = TRUE,
                        build_vignettes = TRUE
)
library(McMasterPandemic)
library(ggplot2); theme_set(theme_bw())
library(cowplot)

@@

<<6: MacPan Params , echo=F>>=
params <- read_params("ICU1.csv")

@@

<<Ali:SIR , echo=F>>=
# A simple SIR model
library(deSolve)
unpack <- McMasterPandemic::unpack

@@

<<Try MacPan to recover the SIR result, echo=FALSE>>=
# update MacPan's params according to the SIR params
params <- update(params, c(beta0=sir.params[["beta"]],
                           alpha=0, ## no asymptomatic
                           sigma=3, ## very short latent period
                           gamma_a=sir.params[["gamma"]],
                           gamma_m=sir.params[["gamma"]],
                           gamma_s=sir.params[["gamma"]],
                           gamma_p=sir.params[["gamma"]]))

@@

gg01 <- (ggplot(sim_combined,aes(x=date))
+ geom_point(aes(y=I))
+ geom_point(aes(y=sir.I), color='red')
)
print(gg01)
print(gg01 + scale_y_log10())
@

>>>>

<<install the package, echo=T  >>=
# install remotes package if necessary:
while (!require(remotes)) {
    install.packages("remotes")
}
## install development version of bbmle:
if (!require("bbmle") || packageVersion("bbmle") < "1.0.23.5") {
    remotes::install_github("bbolker/bbmle")
}
## install the target package and all its dependencies:
remotes::install_github("bbolker/McMasterPandemic",
                        dependencies = TRUE,
                        build_vignettes = TRUE
)
library(McMasterPandemic)
library(ggplot2); theme_set(theme_bw())
library(cowplot)

@@

<<6: MacPan Params , echo=T>>=
params <- read_params("ICU1.csv")

@@

<<Ali:SIR , echo=T>>=
# A simple SIR model
library(deSolve)
unpack <- McMasterPandemic::unpack

@@

<<Try MacPan to recover the SIR result, echo=T>>=
## update MacPan's params according to the SIR params
##
big_rate <- 20
params <- update(params, c(beta0=sir.params[["beta"]],
                           alpha=0, ## no asymptomatic
                           sigma=big_rate, ## very short latent period
                           ## would like this to -> Inf but can't
                           gamma_a=0,
                           mu=1,  ## all cases mild
                           gamma_m=sir.params[["gamma"]],
                           gamma_s=0,
                           ## try to skip presymptomatic period
                           gamma_p=big_rate
                           ))

@@

gg01 <- (ggplot(sim_combined,aes(x=date))
+ geom_point(aes(y=I))
+ geom_line(aes(y=sir.I), color='red')
)
print(gg01)
print(gg01 + scale_y_log10())

### 2020-12-30 note/ProofOfRho.tex

\maketitle
[Fady: for the proof below, it's probably good to include the matrix $V^{-1}$ in the proof. You have it on slide 39/49 from your talk today.]
\begin{theorem}
$\Rnum$ is a decreasing function of $\rho$
\begin{proof}
\ali{Test}We know that $FV^{-1}$ is given by

@@

notice that the columns of $V^{-1}$ always sum to $\gamma$. Furthermore, notice that the nonzero entries of the first row of $V^{-1}$ are strictly decreasing with respect to $\rho$.  Hence, the effect of increasing $\rho$ is to ''shift weight" from the entries in the first row to entries in lower rows. Now recall that

@@

Since $\eta_w < 1$, $\eta_t < 1$, we know that $S_u$ is the largest entry in the first row of $F$, and $S_n$ is the largest entry in the second row of $F$. Thus, whenever $\rho$ increases, the dot product between row vectors of $F$ and column vectors of $V^{-1}$ decreases and hence entries of $FV^{-1}$ decrease. So in particular, the trace of $G_{11}$ (and hence $\Rnum$) decreases.

>>>>

\maketitle
[Fady: for the proof below, it's probably good to include the matrix $V^{-1}$ in the proof. You have it on slide 39/49 from your talk today.]
\begin{theorem}
$\Rnum$ is a decreasing function of $\rho$
\begin{proof}
We know that $FV^{-1}$ is given by

@@

notice that the columns of $V^{-1}$ always sum to $\gamma$ \ali{$1/\gamma$}. Furthermore, notice that the nonzero entries of the first row of $V^{-1}$ are strictly decreasing with respect to $\rho$.  Hence, the effect of increasing $\rho$ is to ''shift weight" from the entries in the first row to entries in lower rows. Now recall that

@@

Since $\eta_w < 1$, $\eta_t < 1$, we know that $S_u$ is the largest entry in the first row of $F$, and $S_n$ is the largest entry in the second row of $F$. Thus, whenever $\rho$ increases, \ali{the next sentence sounds odd} the dot product between row vectors of $F$ and column vectors of $V^{-1}$ decreases and hence entries of $FV^{-1}$ decrease. So in particular, the trace of $G_{11}$ (and hence $\Rnum$) decreases.
